\chapter{Command Line}

\section{Basic commands}

\textit{Covered in \href{https://www.youtube.com/watch?v=oIsH0V3fRt8&list=PLjGmdnqrOKuYXiu7lgG5HW71jPEUd1XCm&index=2}{video 1 of the series}}

\subsection{\texttt{ls}}

\texttt{ls} stands for list. Lists out the files and folders you have in your current directory. \textbf{Directory is another word for folder.}

\begin{lstlisting}[language=bash]
$ ls
README.md   docs  node_modules  package.json    
app     gulpfile.js     package-lock.json
\end{lstlisting}

You could run \texttt{ls -a} to show hidden files and folders as well, which are the ones with their names start with a \texttt{.}

\begin{lstlisting}[language=bash]
$ ls -a
.   .git    docs    package.json
..  .gitignore  gulpfile.js
.DS_Store   README.md   node_modules
.eslintrc   app package-lock.json
\end{lstlisting}

As you can see a hidden folder \texttt{.git} and a hidden file \texttt{.gitignore} are displayed.

\subsection{\texttt{cd}}

\texttt{cd} stands for change directory, self explanatory ;). Use \texttt{cd ..} to go back to the previous directory (to go out one folder). 

\begin{lstlisting}[language=bash]
# KidProf in ~
$ cd code

# KidProf in ~/code
$ cd ..

# KidProf in ~
\end{lstlisting}

\subsection{\texttt{pwd}}

\texttt{pwd} shows the current directory you are in in full. As shown in chapter1 it makes it easier for you to locate our work folder in Finder/ File Explorer.

\begin{lstlisting}[language=bash]
# KidProf in ~
$ pwd
/Users/KidProf
\end{lstlisting}

\subsection{\texttt{touch}}

Creates a new file. For example, \texttt{touch test.txt}

\subsection{\texttt{mkdir}}

Creates a new folder. For example, \texttt{mkdir code}

\subsection{\texttt{--version}}

To check the version of a certain software. It is commonly used to verify that you have installed the software correctly. 

\begin{lstlisting}[language=bash]
(node is installed)
$ node --version
v16.14.0

(node is not installed)
$ node --version
bash: command not found: node
\end{lstlisting}

\section{A word on directories}

\texttt{.} refers to the current directory.

\texttt{..} refers to the previous directory.

\texttt{$\sim$} refers to the home directory, it is the initial directory when you open the command line. In the \texttt{pwd} example, you can see the home directory of my Mac is \texttt{/Users/KidProf}. You should do most of your work inside this home directory. To go back to the home directory, do \texttt{cd $\sim$} This is useful when you get stuck in some other directories.

When the path is not prepended with a \texttt{/}, it is a relative path. Meaning that it is relative to the current directory.

When the path is prepended with a \texttt{/}, it is an absolute path. 

\begin{lstlisting}[language=bash]
(cd to an relative path)
# KidProf in ~
$ cd code

# KidProf in ~/code

(cd to an absolute path)
# KidProf in ~
$ cd /Users/KidProf/code

# KidProf in ~/code
\end{lstlisting}

\section{General Tips}

\textit{Covered in \href{https://www.youtube.com/watch?v=oIsH0V3fRt8&list=PLjGmdnqrOKuYXiu7lgG5HW71jPEUd1XCm&index=2}{video 1 of the series}}
\vspace{6mm}

Use up and down arrow to navigate your command history. Very useful when you are calling the same command repeatedly or when you made some typos.

Press \texttt{ctrl+C} to .... % TODO

\section{More commands}

\textit{Advanced, }
\textit{covered in \href{https://www.youtube.com/watch?v=ZIBEVGrtiVA&list=PLjGmdnqrOKuYXiu7lgG5HW71jPEUd1XCm&index=7}{video 6a of the series}}
\vspace{6mm}

\subsection{\texttt{cp}}

\subsection{\texttt{rm}}

\subsection{\texttt{mv}}

\section{Vim - command line text editor}

\textit{Advanced}
\vspace{6mm}

The default text editor is called Vim. It may appear when you run some commands that require messages to be inputted, for example, \texttt{git commit} \textit{(More in Chapter 3)}.

Unfortunately, it is quite difficult to use and needs some time to get used to. So I tried my best to provide alternative commands that prevent this text editor from popping up, for example, using \texttt{git commit -m "message"}

However, here are the basics of using Vim, so that you will not get too lost when you encounter it. You can practice by entering \texttt{vim} in the command line.

You start in a mode called "normal mode". You can’t immediately type anything.

In order to get typing press \texttt{i} (stands for insert). This will bring you to "insert mode", so named because in this mode you can actually type.

When you are done typing press \texttt{esc}. This will bring you back to "normal mode".

In order to save your work you want to type :w and press return. And in order to exit vim you want to type \texttt{:q} and press return. Because saving and quitting is a very common action, there is actually a shortcut \texttt{:x}, which stands for \texttt{:wq} (which just combines \texttt{:w} and \texttt{:q}).\footnote{Reference: \url{https://web.mit.edu/6.005/www/fa14/tutorial/git/config.html}}


If you do not wish to save the file, you can use \texttt{:q!}


